Joe Biden Predicted to Win 55.9% of the Popular Vote in the 2024 U.S. Presidential Election

### 2024-03-17 paper/paper.qmd
updated paper

@@

@tbl-popvotepost-popularvote summarizes the popular vote and electoral college predictions

The estimates for the support for Biden in each state are shown in @tbl-estimatedpropsupport. We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport

@@

@@

## Weaknesses and Limitations {#sec-weaknesses-limitations}

@@

## Next Steps

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

@@

@fig-stanareyouokay-2 shows the trace plot for our variables. We can see that the lines bounce, but remain horizontal and have a lot of overlap with each other [@tswd]. This suggests that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors.

@@

plot(presidential_vote_model, "trace")

@@

In @fig-modelplot, we show the 90% credibility intervals for the predictors of `vote_biden`.

>>>>

We got our popular vote prediction by taking the weighted average of vote_biden across all U.S. states and converting it into a percent. The average is weighted according to the number of census respondents in our post-stratification data set from each state. @tbl-popvotepost-popularvote summarizes our mean, lower, and upper estimates for the popular vote percentage and number of electoral college votes that Joe Biden will receive on November 5th. Based on the results of our multi-level regression with post-stratification (MRP), which involved applying the logistic regression outlined in @sec-model to our post-stratification data set, we predict that Joe Biden and Donald Trump will receive 56.59% and 43.41%, respectively, of the popular vote. That means that our logistic regression model classified 56.59% of the people included in our subset of the ACS 2022 [@citeIPUMS] as Biden voters based on sex, age bracket, race, highest level of education, and whether they live in an urban or rural area. @tbl-popvotepost-popularvote shows our lower, mean, and upper estimate for the percent of the popular vote that Joe Biden will receive. Conversely, the lower estimate for Joe Biden corresponds to the upper estimate for Donald Trump, and the upper estimate for Joe Biden

@@

The estimates for the support for Biden in each state are shown in @fig-estimatedpropsupport. We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport. We can see that the proportion of voters classified as Biden voters

@@

## Electoral College Prediction

@@

## Weaknesses, Limitations, and Possible Next Steps {#sec-weaknesses-limitations}

@@

With a survey data set that is as limited as the one we currently have, SoftMax regression is not a viable option. However, the possibility of third-party candidates receiving vote shares larger than the margin of victory of the Democratic or Republican nominee in key battleground states highlights the need to consider the fact that a small percentage of voters will cast a ballot for someone not named Donald Trump or Joe Biden on November 5th. The first step towards extending and improving this report in the future would be to find a survey data set that has significantly more respondents, is somewhat reflective of voting trends and demographics in each state, and specifically asks respondents who they plan to vote for in the 2024 U.S. presidential election. Once we address the issue of having very few, or even no respondents from smaller states and no question about 2024 vote intention, we will be able to build a stronger and more accurate model to predict vote choice based on state, urban or rural area, demographics such as sex, age, and race, and education. A robust survey data set with enough respondents (possibly upwards of 10,000) will allow us to train a more powerful SoftMax regression model to perform multi-class classification and classify voters as Trump, Biden, third-party, or abstaining voters using similar predictors to the ones outlined in @sec-model as well as state.

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.

@@

@fig-stanareyouokay-2 shows the trace plot for our intercept. We can see that the lines bounce, but remain horizontal and have a lot of overlap with each other [@tswd]. @fig-tracesex, @fig-traceage, @fig-tracerace, @fig-traceeducation, and @fig-traceurban show the trace plots for our sex, age_bracket, races, education_level, and urban variables. Each of these trace plots looks similar to @fig-stanareyouokay-2, as the lines bounce, remain horizontal, and overlap a lot. Our trace plots suggest that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors [@tswd].

@@

plot(presidential_vote_model, "trace","(Intercept)")

@@

In @fig-modelplot, we show the 90% credibility intervals for the predictors of `vote_biden`. We see that the 90% credibility interval for when urban=urban is greater than 0. This means that a respondent who lives in an urban area is more likely to support Joe Biden. We also see that the 90% credibility interval for sex=male is below zero. This means that based on our survey data and model, male respondents are more likely to support Donald Trump.
